Problem Set 1 (Answers)

### 2019-09-26 content/class/09-class.Rmd
Nicer formatting to problem set 1 answer key html, and link to it in class pages

[Problem Set 1](/assignment/01-problem-set.html) answers will be posted shortly, and the graded assignment emailed back to you.^[Sorry I've been too busy to send emails confirming that I recieved your problem set!]

>>>>

[Problem Set 1](/assignment/01-problem-set.html) answers are posted, and the graded assignment emailed back to you.^[Sorry I've been too busy to send emails confirming that I recieved your problem set!]

### 2019-09-26 static/homeworks/01-problem-set-answers.Rmd
Nicer formatting to problem set 1 answer key html, and link to it in class pages

## 1. 
### a. 
**What are the top 5 boys names for 2017, and what *percent* of overall names is each?**

@@

### b.
**What are the top 5 *girls* names, and what *percent* of overall names is each?**

@@

## 2. 
**Make two barplots, of these top 5 names, one for each sex. Map `aes`thetics `x` to `name` and `y` to `prop`^[Or `percent`, if you made that variable, as I did.] and use `geom_col` (since you are declaring a specific `y`, otherwise you could just use `geom_bar()` and just an `x`.)**

@@

## 3. 
**Find your name.^[If your name isn't in there :(, pick a random name.] `count` by `sex` how many babies since 1880 were named your name.^[Hint: if you do this, you'll get the number of *rows* (years) there are in the data. You want to add the number of babies in each row (`n`), so inside `count`, add `wt=n` to weight the count by `n`.] Also add a variable for the percent of each sex.**

@@

## 4.
**Make a line graph of the number of babies with your name over time, `color`ed by `sex`.**
```{r}
babynames %>%
  filter(name == "Ryan") %>%
  ggplot(data = .)+
  aes(x = year,
      y = n,
      color = sex)+
  geom_line(size=1)+
  labs(x = "Year",
       y = "Number of Babies",
       title = "Popularity of Babies Named 'Ryan'",
       color = "Sex",
       caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)
```

## 5.
### a.
**Make a table of the most common name for boys by year between 1980-2017.^[Hint: once you've got all the right conditions, you'll get a table with a lot of data. You only want to `slice` the `1`st row for each table.]**

@@

### b.
**Now do the same for girls.**

@@

## 6. Now let's graph the evolution of the most common names since 1880.
### a.
**First, find out what are the top 10 *overall* most popular names for boys and for girls. You may want to create two vectors, each with these top 5 names.**

@@

### b.
**Now make two `line`graphs of these 5 names over time, one for boys, and one for girls.**

@@

## 7.
**_Bonus (hard!)_: What are the 10 most common "gender-neutral" names?^[This is hard to define. For our purposes, let's define this as names where between 48 and 52% of the babies with the name are Male.]**

@@

---

@@

## 8. 
**Download these two datasets that I've cleaned up a bit:^[If you want, try downloading them from the websites yourself!]**

@@

## 9. 
**The `polfreedom` dataset is still a bit messy. Let's overwrite it (or assign to something like `polfreedom2`) and select `Country/Territory` and `Total` (total freedom score) and rename `Country.Territory` to `Country`.**

@@

## 10.
**Now we can try to merge these two datasets into one. Since they both have `Country` as a variable, we can merge these tibbles using `left_join(econfreedom, polfreedom, by="Country")`^[Note, if you saved as something else in question 9., use that instead of `polfreedom`!] and save this as a new tibble (something like `freedom`).**

@@

## 11.
**Now make a scatterplot of Political Freedom (`total`)^[Feel free to `rename` these!] as `y` on Economic Freedom (`ef`) as `x` and `color` by `continent`.**

@@

## 12.
**Let's do this again, but highlight some key countries. Pick three countries, and make a new tibble from `freedom` that is only the observations of those countries. Additionally, *install* and *load* a packaged called `ggrepel`^[This automatically adjusts labels so they don't cover points on a plot!] Next, redo your plot from question 11, but now add a layer: `geom_label_repel` and set its `data` to your three-country tibble, use same `aes`thetics as your overall plot, but be sure to add `label = ISO`, to use the ISO country code to label.^[You might also want to set a low `alpha` level to make sure the labels don't obscure other points!]**

@@

## 13.
**Make another plot similar to 12, except this time use GDP per Capita (`gdp`) as `y`. Feel free to try to put a regression line with `geom_smooth()`!**^[If you do, be sure to set its data to the full `freedom`, not just your three countries!]. **Those of you in my Development course, you just made my graphs from [Lesson 2](https://devf19.classes.ryansafner.com/slides/02-slides#23)!**

>>>>

## Question 1

### Part A 
**What are the top 5 boys names for 2017, and what *percent* of overall names is each?**

@@

### Part B
**What are the top 5 *girls* names, and what *percent* of overall names is each?**

@@

## Question 2
**Make two barplots, of these top 5 names, one for each sex. Map `aes`thetics `x` to `name` and `y` to `prop`^[Or `percent`, if you made that variable, as I did.] and use `geom_col` (since you are declaring a specific `y`, otherwise you could just use `geom_bar()` and just an `x`.)**

@@

## Question 3
**Find your name.^[If your name isn't in there :(, pick a random name.] `count` by `sex` how many babies since 1880 were named your name.^[Hint: if you do this, you'll get the number of *rows* (years) there are in the data. You want to add the number of babies in each row (`n`), so inside `count`, add `wt=n` to weight the count by `n`.] Also add a variable for the percent of each sex.**

@@

## Question 4
**Make a line graph of the number of babies with your name over time, `color`ed by `sex`.**
```{r}
babynames %>%
  filter(name == "Ryan") %>%
  ggplot(data = .)+
  aes(x = year,
      y = n,
      color = sex)+
  geom_line(size=1)+
  labs(x = "Year",
       y = "Number of Babies",
       title = "Popularity of Babies Named 'Ryan'",
       color = "Sex",
       caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)
```

## Question 5
### Part A
**Make a table of the most common name for boys by year between 1980-2017.^[Hint: once you've got all the right conditions, you'll get a table with a lot of data. You only want to `slice` the `1`st row for each table.]**

@@

### Part B
**Now do the same for girls.**

@@

## Question 6
**Now let's graph the evolution of the most common names since 1880.**
### Part A
**First, find out what are the top 10 *overall* most popular names for boys and for girls. You may want to create two vectors, each with these top 5 names.**

@@

### Part B
**Now make two `line`graphs of these 5 names over time, one for boys, and one for girls.**

@@

## Question 7
**_Bonus (hard!)_: What are the 10 most common "gender-neutral" names?^[This is hard to define. For our purposes, let's define this as names where between 48 and 52% of the babies with the name are Male.]**

@@

@@

## Question 8
**Download these two datasets that I've cleaned up a bit:^[If you want, try downloading them from the websites yourself!]**

@@

## Question 9 
**The `polfreedom` dataset is still a bit messy. Let's overwrite it (or assign to something like `polfreedom2`) and select `Country/Territory` and `Total` (total freedom score) and rename `Country.Territory` to `Country`.**

@@

## Question 10
**Now we can try to merge these two datasets into one. Since they both have `Country` as a variable, we can merge these tibbles using `left_join(econfreedom, polfreedom, by="Country")`^[Note, if you saved as something else in question 9., use that instead of `polfreedom`!] and save this as a new tibble (something like `freedom`).**

@@

## Question 11
**Now make a scatterplot of Political Freedom (`total`)^[Feel free to `rename` these!] as `y` on Economic Freedom (`ef`) as `x` and `color` by `continent`.**

@@

## Question 12
**Let's do this again, but highlight some key countries. Pick three countries, and make a new tibble from `freedom` that is only the observations of those countries. Additionally, *install* and *load* a packaged called `ggrepel`^[This automatically adjusts labels so they don't cover points on a plot!] Next, redo your plot from question 11, but now add a layer: `geom_label_repel` and set its `data` to your three-country tibble, use same `aes`thetics as your overall plot, but be sure to add `label = ISO`, to use the ISO country code to label.^[You might also want to set a low `alpha` level to make sure the labels don't obscure other points!]**

@@

## Question 13
**Make another plot similar to 12, except this time use GDP per Capita (`gdp`) as `y`. Feel free to try to put a regression line with `geom_smooth()`!**^[If you do, be sure to set its data to the full `freedom`, not just your three countries!]. **Those of you in my Development course, you just made my graphs from [Lesson 2](https://devf19.classes.ryansafner.com/slides/02-slides#23)!**
